Novel Sampling Using Stochastic Gradient Descent

### 2016-12-07 paper/main.tex

$f_1(x) = \exp(-x^4)$

$f_2(x) = \exp(-(x^2 - 4)(x^2 - 1))$

$f_3(x) = \exp(-|x|)$

- what we have achieved

- Future work

>>>>

In this paper, we have shown that stochastic gradient descent approaches extend in an elegant way to a sampler for posterior distributions when computing the likelihood from the full data is intensive. We are able to establish the stationary distribution of the procedure and gave several examples showing the accuracy and uses of our result. By taking a Bayesian route, we are able to show that the SGD algorithm is a sampler of a posterior under a non-informative prior. This enables us to establish connection between SGD and maximum \emph{a posteriori} (MAP) procedure, which is an analogy to gradient descent solving a maximum likelihood estimator in the frequentist paradigm. A natural extension of this work, which we hope to be able to accomplish, is to derive finite sample rates of convergence for the mixing time, which would allow a more rigor understanding of the SGD family as a whole. Extensive simulation of the methodology is also needed to fully understand this novel SGD sampler.

### 2016-12-07 poster/main.tex

% inline theorem
\makeatletter
\setbeamertemplate{theorem begin}
  {\usebeamercolor[fg]{thcolor}% for the heading
  {\bfseries\inserttheoremname~}%
  \ifx\inserttheoremaddition\@empty\else(\inserttheoremaddition)\ \fi%
  \hspace{.01em}\normalfont\usebeamercolor[fg]{thcolor}% for the body
  }
\setbeamertemplate{theorem end}{}
\makeatother
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

@@

\begin{block}{SGD As Fast Posterior Sampler}
When $p(\theta)$ is the posterior distribution of some parameters given a set of data $X_1,\ldots, X_n$, i.e.,
\[
p(\theta) = \prod_{i=1}^n P(X_i|\theta) P(\theta),
\]
The noisy gradient step in Algorithm 1 can be estimated by subsampling:
(Central limit theorem)
Let $K$ be the subsampling size and $\eta_1,\ldots, \eta_K$ be samples drawing from $1,\ldots, n$ with replacement, then
\[
K^{-1}\sum_{i=1}^K (\nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)) \sim \N(n^{-1} \nabla \log p(\theta) , K^{-1}\cov \nabla \log p(\theta)).
\]

@@

\begin{block}{Upper Bound On Mixing Time}
How effective Algorithm 1 is as a sampling method? Denote $\|\cdot\|_{\mathrm{TV}}$ the TV norm and $P^t(\bx^0)$ the density of $\bx^t$ starting from $\bx^0$. Define its mixing time to be
\[
\tau_{\mathrm{mix}} \triangleq \min\{t|\max_{\bx^0}\|P^t(\bx^0) - p\|_{\mathrm{TV}}\leq 1/4\}.
\]
Suppose $\tau_{\bx^0}$ is the stopping time when the algorithm converges to global minimum starting from $\bx^0$, we might be able to have a upper bound for $\tau_{\mathrm{mix}}$.
\[
\tau_{\mathrm{mix}} \leq 4 \max_{\bx^0}\E \tau_{\bx^0}.
\]
So when SGD converges fast, it will also be efficient as a sampling method.
\end{block}

@@

\begin{itemize}
\item Proposed fast SGD sampler for posterior distributions when computing the likelihood using full data is intensive
\item Proved accurate stationary distribution based on oracle results
\item Established connection between SGD and maximize \emph{a posteriori}
\item Derived upper bound on mixing time
\item Future work: Finite sample rates of convergence for the mixing time
\end{itemize}

>>>>

% inline theorem
% \makeatletter
% \setbeamertemplate{theorem begin}
%   {\usebeamercolor[fg]{thcolor}% for the heading
%   {\bfseries\inserttheoremname~}%
%   \ifx\inserttheoremaddition\@empty\else(\inserttheoremaddition)\ \fi%
%   \hspace{.01em}\normalfont\usebeamercolor[fg]{thcolor}% for the body
%   }
% \setbeamertemplate{theorem end}{}
% \makeatother
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

@@

\begin{block}{SGD As Fast Posterior Sampler}
When $p(\theta)$ is the posterior distribution of some parameters given a set of data $X_1,\ldots, X_n$, i.e.,
$$
p(\theta) = \prod_{i=1}^n P(X_i|\theta) P(\theta),
$$
The noisy gradient step in Algorithm 1 can be estimated by subsampling:

(Central limit theorem)
Let $K$ be the subsampling size and $\eta_1,\ldots, \eta_K$ be samples drawing from $1,\ldots, n$ with replacement, then
\[
K^{-1}\sum_{i=1}^K (\nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)) \sim \N(n^{-1} \nabla \log p(\theta) , K^{-1}\cov \nabla \log p(\theta)).
\]

@@

\begin{block}{Upper Bound On Mixing Time}
How effective is Algorithm 1 as a sampling method? Denote $\|\cdot\|_{\mathrm{TV}}$ the TV norm and $P^t(\bx^0)$ the density of $\bx^t$ starting from $\bx^0$. Define its mixing time to be
\[
\tau_{\mathrm{mix}} \triangleq \min\{t|\max_{\bx^0}\|P^t(\bx^0) - p\|_{\mathrm{TV}}\leq 1/4\}.
\]
Suppose $\tau_{\bx^0}$ is the stopping time when the algorithm converges to the global minimum starting from $\bx^0$, we might be able to have an upper bound for $\tau_{\mathrm{mix}}$.
\[
\tau_{\mathrm{mix}} \leq 4 \max_{\bx^0}\E \tau_{\bx^0}.
\]
So when SGD converges fast, it will also be efficient as a sampling method.
\end{block}

@@

\begin{itemize}
\item Proposed fast SGD sampler for posterior distributions when computing the likelihood using full data is intensive
\item Proved accurate stationary distribution based on oracle results
\item Established connection between SGD and maximum \emph{a posteriori}
\item Derived upper bound on mixing time
\item Future work: Finite sample rates of convergence for the mixing time
\end{itemize}
